Excerno analysis

### 2022-08-05 paper_analysis.Rmd
Updated figure 3 for dillution analysis

```{r}
(roc_plot <- 
  ggplot(allsim_tbl, aes(d=truth, m=bayes, color=dataset))+
  geom_roc(cutoffs.at = c(0.5), cutoff.labels = c(0.5)) +
  labs(x="False Positive Rate",
       y="True Positive Rate",
       color="Signature"))

ggsave("paper_figures/fig2a.png", roc_plot, width = 131.25, height = 108, units = "mm")

@@

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tp + fn),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

@@

```{r}
plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
```

@@

>>>>

```{r}
scale = seq (0, 1, by=0.2)

(roc_plot <- 
  ggplot(allsim_tbl, aes(d=truth, m=bayes, color=dataset))+
  geom_roc(cutoffs.at = c(0.5), cutoff.labels = c(0.5)) +
  labs(x="False Positive Rate",
       y="True Positive Rate",
       color="Signature")+
  scale_x_continuous(labels=scale, breaks=scale)+
  scale_y_continuous(labels=scale, breaks=scale))

ggsave("paper_figures/fig2a.png", roc_plot, width = 150, height = 90, units = "mm")

@@

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tn + fp),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

@@

```{r}

plot_metrics_excerno(sim_tbl, "Accuracy", 0.8) +
  scale_x_continuous(labels=scale, breaks=scale)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)+
    scale_x_continuous(labels=scale, breaks=scale)
ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)+
  scale_x_continuous(labels=scale, breaks=scale)
ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
```

@@

# What is the influence of FFPE percentage in the performance of `excerno`?

We are interested in the performance characteristics of excerno under different percentages of FFPE signature.

## An initial simulation

So we will mimick our original scenario scenario of SBS4 and SBS6, however this time we will introduce varying proportions going from 5% to 100% every 5%.

Let's start with SBS4

```{r}
set.seed(654321)
pct_ffpe = seq (10,90, by=5)
metrics_order = c("Sensitivity","Specificity", "Accuracy")

sbs4_dilution_tbl <- pct_ffpe %>%
  map_dfr(~calc_metrics_mix_num(4,10, (100-.x)*10, .x*10)) %>%
  mutate(pct_ffpe = rep(pct_ffpe, each=3),
         name = factor (name, levels = metrics_order))
```

And plot it

```{r}
ggplot(sbs4_dilution_tbl, 
       aes(x=pct_ffpe, y=mean)) +
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(name~.) + 
  geom_hline(yintercept=0.8, linetype="dashed",color = "red")
```

Seems the metrics stay above 80%, except for specificity at 10% which slightly dips below 80%

Let's do the same analysis with SBS6:

```{r}
sbs6_dilution_tbl <- pct_ffpe %>%
  map_dfr(~calc_metrics_mix_num(6,10, (100-.x)*10, .x*10)) %>%
  mutate(pct_ffpe = rep(pct_ffpe, each=3),
         name = factor (name, levels = metrics_order))

ggplot(sbs6_dilution_tbl, 
       aes(x=pct_ffpe, y=mean)) +
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(name~.) + 
  geom_hline(yintercept=0.8, linetype="dashed",color = "red")
```

Here we see how sensitivity dips below our threshold for all ffpe percentages *above* 60% ffpe and specificity drops below 80% for all the % ffpe *below* 50%. Accuracy stays mostly above 80% and it is best when there is only one class of mutations (either FFPE or a single SBS)

Let's combine both datasets in a single graph by using color and focus on sensitivity and specificity.

```{r}
dilution_tbl <- rbind (sbs4_dilution_tbl, sbs6_dilution_tbl)

scale_x = seq (10,90, by=10)
scale_y = seq (0.2, 1, by=0.2)

dilution_tbl %>%
  filter (name!="Accuracy") %>%
  ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
    geom_point()+
    geom_smooth(se=FALSE)+
    facet_grid(name~.) + 
    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
    labs(x= "Percent of FFPE signature",
         y="",
         color = "Signature")+
    scale_x_continuous(breaks = scale_x, labels = scale_x)+
    scale_y_continuous(breaks = scale_y, labels = scale_y)

ggsave("paper_figures/fig3a.png", dpi = 300, width = 150, height = 90, units = "mm")
```

Notice the trend is that for increasing percent of FFPE signature, the sensitivity decreases, while the specificity increases.

## A general approach

We are interested in assessing the trends in performance metrics across *all* of the signatures in COSMIC. In order to do that we need to create an auxiliary function `calc_dillution_metrics` that generates a tibble with this information for a particular mutational signature. We also create a function `plot_dillution_metrics` that plots this information

```{r}
calc_dillution_metrics <- function (ffpe_grid, cosmic_idx) {
  metrics_order = c("Sensitivity","Specificity", "Accuracy")
  sig_name=factor(signature_names[cosmic_idx], signature_names)

  ffpe_grid %>%
    map_dfr(~calc_metrics_mix_num(cosmic_idx,10, (100-.x)*10, .x*10)) %>%
    mutate(pct_ffpe = rep(ffpe_grid, each=3),
         sbs = sig_name,
         metric = factor (name, levels = metrics_order)) %>%
    select(-c(name))
}

plot_dillution_metrics <- function (sim_tbl) {
  scale_x = seq (10,90, by=10)
  scale_y = seq (0.2, 1, by=0.2)

  sim_tbl %>%
  ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
    geom_point()+
    geom_smooth(se=FALSE)+
    facet_grid(metric~.) + 
    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
    labs(x= "Percent of FFPE signature",
         y="",
         color = "Signature")+
    scale_x_continuous(breaks = scale_x, labels = scale_x)+
    scale_y_continuous(breaks = scale_y, labels = scale_y)
}
```

We test our functions for SBS4 and SBS6:

```{r}
calc_dillution_metrics(pct_ffpe, 4) %>%
  plot_dillution_metrics()

calc_dillution_metrics(pct_ffpe, 6) %>%
  plot_dillution_metrics()
```

Now we are ready to calculate the dillution metrics across all COSMIC mutational signatures (this takes 20min to run)

```{r}
set.seed(123456)
# Full dataset run
sim_dillution_tbl <- 1:sbs_num %>%
  map_dfr(~calc_dillution_metrics(pct_ffpe, .x))
```

And we plot the sensitivity and specificity across all levels of FFPE contamination:

```{r}
scale_x = seq (10,90, by=10)
scale_y = seq (0.2, 1, by=0.2)

sim_dillution_tbl %>%
  filter(metric != "Accuracy") %>%
  ggplot (aes(x=pct_ffpe, y=mean, color=metric)) +
    geom_point (alpha=0.3) +
    geom_smooth(method="lm", se=FALSE)+
    facet_grid(metric~.)+
    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
    scale_x_continuous(breaks = scale_x, labels = scale_x)+
    scale_y_continuous(breaks = scale_y, labels = scale_y)+
    labs(x= "Percent of FFPE signature",
         y="")+
    theme(legend.position = "none")

ggsave("paper_figures/supp5.png", dpi = 300, width = 150, height = 90, units = "mm")
```
